![](https://www.unibg.it/themes/custom/unibg/logo.svg){width="250"}

# [Geomatching]{.underline}: a sneak peek

## Import R libraries

```{r}

rm(list = ls())
library(ggplot2)
library(mapdata)
library(geotools)
```

## Import (local) data

-   `ERA5Land_t2m_10d`: time-spatial data.frame (*longitude*, *latitude* and *time*) containing 10 days of **air temperature**. Grid resolution: 0.1° x 0.1° (about 11 km x 11 km).

-   `ERA5SL_t2m_10d`: time-spatial data.frame containing 10 days of **air temperature**. Grid resolution: 0.25° x 0.25° (about 27 km x 27 km).

-   `ERA5SL_rh_10d`: time-spatial data.frame containing 10 days of **relative humidity**. Grid resolution: 0.25° x 0.25° (about 27 km x 27 km).

-   `ERA5SL_lai_hv_10d_matrix`: 3D matrix (*x*-component represents *longitude, y*-component describes *latitude* and *z-*component represents *time*) 10 days of **Leaf Area Index - High Vegetation**. Grid resolution: 0.25° x 0.25° (about 27 km x 27 km).

```{r}

# weather data
load("demo_data.Rdata")
names(ERA5Land_t2m_10d)[4] <- "t2m_Land"
names(ERA5SL_t2m_10d)[4] <- "t2m_SL"
WE_data = list(
  "ERA5Land_t2m" = list(
    "data" = ERA5Land_t2m_10d,
    "var" = "t2m_Land",
    "label" = "Air temperature, ERA5Land (11 km x 11 km)"
  ),
  "ERA5SL_t2m" = list(
    "data" = ERA5SL_t2m_10d,
    "var" = "t2m_SL",
    "label" = "Air temperature, ERA5SL (27 km x 27 km)"
  ),
  "ERA5SL_rh" = list(
    "data" = ERA5SL_rh_10d,
    "var" = "rh",
    "label" = "Relative humidity, ERA5SL (27 km x 27 km)"
  )
)

# LAUs boundaries
load("../data/IT_adm_bounds_2025.RData")
names(mun_bounds)[6] <- "COD_COM"
LAUs_data = list(
  "mun_bounds" = list(
    "data" = mun_bounds,
    "code" = "mun",
    "join_var" = "COD_COM"
  ),
  "prov_bounds" = list(
    "data" = prov_bounds,
    "code" = "prov",
    "join_var" = "COD_PROV"
  ),
  "reg_bounds" = list(
    "data" = reg_bounds,
    "code" = "reg",
    "join_var" = "COD_REG"
  )
)
```

### Visualize first rows of each data.frame

`ERA5Land_t2m_10d`

```{r}

head(ERA5Land_t2m_10d, n = 10)
```

`ERA5SL_t2m_10d`

```{r}

head(ERA5SL_t2m_10d, n = 10)
```

`ERA5SL_rh_10d`

```{r}

head(ERA5SL_rh_10d, n = 10)
```

## First look

```{r}

for (item_i in WE_data) {
  
  # configuration
  df_i <- item_i$data
  var_i <- item_i$var
  label_i <- item_i$label
  subset_i <- df_i[df_i$time == df_i$time[1], ]
  
  # run
  p <- ggplot() +
    geom_map(
      data = map_data("italy"),
      map = map_data("italy"),
      aes(map_id = region),
      fill = "lightgrey",
      color = "black",
      linewidth = .1,
      alpha=.9
    ) + 
    geom_tile(
      data = subset_i,
      aes(x = longitude, y = latitude, fill = .data[[var_i]])
    ) +
    labs(
      title = sprintf("Geographical map of weather data"),
      subtitle = label_i,
      x = "Longitude",
      y = "Latitude"
    ) + 
    scale_fill_viridis_c(
      option = "C"
    ) +
    theme(
      legend.title = element_blank(),
      plot.title = element_text(face = "bold")
    )
  print(p)
}
```

## Run geomatching

```{r}
aggr_data <- list()
for (item_i in LAUs_data) {
  
  # configuration
  bounds_i <- item_i$data
  code_i <- item_i$code
  join_var_i <- item_i$join_var
  
  # run
  temp <- geotools::geomatching(
    data = list(
      ERA5Land_t2m_10d,
      ERA5SL_t2m_10d,
      ERA5SL_rh_10d,
      ERA5SL_lai_hv_10d_matrix
    ),
    settings = list(
      "format" = list(
        "xyt",
        "xyt",
        "xyt",
        "matrix"
      ),
      "type" = list(
        "grid",
        "grid",
        "grid",
        "grid"
      ),
      "crs" = list(
        4326,
        4326,
        4326,
        4326
      )
    ),
    aggregate = TRUE,
    group_by = code_i
  )
  
  # rename columns #1
  names(temp)[24:30] <- c(
    "lai_hv_min",
    "lai_hv_1st_quartile",
    "lai_hv_mean",
    "lai_hv_median",
    "lai_hv_3rd_quartile",
    "lai_hv_max",
    "lai_hv_std"
  )
  
  # fill missing values
  stats <- c("min", "1st_quartile", "mean", "median", "3rd_quartile", "max", "std")
  for (stat_i in stats) {
    land_col <- paste0("t2m_Land_", stat_i)
    sl_col <- paste0("t2m_SL_", stat_i)
    temp[[land_col]][is.na(temp[[land_col]])] <- temp[[sl_col]][is.na(temp[[land_col]])]
  }
  
  # drop columns
  temp <- temp[, !names(temp) %in% c(
    "t2m_SL_min", "t2m_SL_1st_quartile", "t2m_SL_mean", "t2m_SL_median",
    "t2m_SL_3rd_quartile", "t2m_SL_max", "t2m_SL_std")
  ]
  
  # rename columns #2
  names(temp)[3:9] <- c(
    "t2m_min",
    "t2m_1st_quartile",
    "t2m_mean",
    "t2m_median",
    "t2m_3rd_quartile",
    "t2m_max",
    "t2m_std"
  )
  
  # join
  temp <- left_join(
    bounds_i,
    temp,
    by = join_var_i
  )
  
  # save
  aggr_data[[code_i]] <- temp
}
```

## Export aggregated data

```{r}

save(aggr_data, file = "WE_aggr_data.RData")
```

## Visualize aggregated data

```{r}

ggplot() + 
  geom_sf(data = aggr_data[["mun"]], aes(fill = t2m_mean), linewidth = .01) + 
    scale_fill_continuous(type = "viridis")

```
